Trends of SDR in Thessaloniki

### 2023-11-04 MDPI_submition.Rmd
Commit 2023-11-04 14:14

In order to investigate further the features of the CUSUMs, we created another set of CUSUM plots by subtracting the corresponding long term trend from the SDR anomaly data, prior to the CUSUM calculation (Figure\ \ref{fig:cusumnotrendmonthly}).
With this approach periods when the CUSUMs diverge from zero can be interpreted as a systematic variation of SDR from the climatological mean. 
When the CUSUM is increasing, the added values are above the climatological values of the SDR trend and vice versa.
Overall, for all- and clear-sky conditions (Figures\ \ref{fig:cusumnotrendmonthly-1} and\ \ref{fig:cusumnotrendmonthly-2}) we observe periods when the anomalies diverge from the climatological value, each lasting for several years.
The pattern in both datasets is very similar, suggesting prevalence in clear skies over Thessaloniki.
It is interesting that in the period 1993 - 2016 the anomalies have a high variability around zero. After 2016, the range of the variability is decreased at about one third of the prior period.
For cloud-sky conditions (Figure\ \ref{fig:cusumnotrendmonthly-3}) the period 1997 - 2008 is dominated by positive CUSUMs suggesting a reduced effect of clouds on SDR.
From 1997 to mid-2000s CUSUMs are increasing, likely due to a continuous decrease in the thickness of clouds, followed by a period of rapid increase (within 3 years) in cloud optical thickness lasting up to 2008.
The following stable period spans for about 15 years up to 2021 when CUSUMs start increasing again.

@@

Similarly to the long term trends discussed above, we have calculated the trend of the anomalies for the three different sky conditions, and for each season of the year, using the corresponding mean monthly values
(Figure\ \ref{fig:seasonalALL} and Table\ \ref{tab:trendseasontable}).
For all-sky conditions the trend in SDR in winter is the largest
($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$),
followed by the trend in autumn
($`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$,
a value close to the long term trend) both statistically significant above the $99\,\%$ confidence level.
In spring and summer, the trends are much smaller and of lesser statistical significance.
These seasonal differences indicate a possible relation of the trends in SDR to trends of clouds during winter and autumn.
For clear-skies, the trend in winter is $`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear sky"], 2)`\,\%/y$, larger than for all-skies ($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$), which is another indication of increasing cloud thickness.
Moreover, the trends under clear- and cloud-sky conditions are almost complementary to each other, particularly for winter and autumn, where the signal is stronger.
During spring and summer the statistical significance is very low and the actual trend too small for a meaningful comparison.

```{r seasonalALL, echo=F, fig.cap="SDR trends by season of the year for the three sky conditions. Each point is a monthly mean calculated from daily means. The vertical scale is common for all graphs, each column of graphs corresponds to a sky condition and each row to a season of the year.", out.width="100%"}
knitr::include_graphics("./images/SeasonalTrendsTogether3-2.png")
```

@@

We demonstrate that in the period
`r paste(year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)))`, there is a brightening trend of SDR for all-sky conditions ($`r alltrend`\,\%/y$).
On a previous study [@Bais2013], for the period 1993 - 2011 the trend was found to be $0.33\,\%/y$.
The increase of this trend, indicates that the phenomenon and probably the causes, are still occurring.
Accordingly, we found a similar trend for clear-sky conditions ($`r cleartrend`\,\%/y$) that further supports that the brightening can be independent of the cloud occurrences, although, cloud formation can also attenuate SDR.
We have calculated the effect of TSI variability to be a negligible factor in SDR variability.
Unfortunately, there is no available data for this period for the aerosols, in order to directly estimate their effect on SDR, which is the most probable main cause.
The attenuation of SDR by aerosols over Europe have been proposed as major factor by @Wild2021.
The dimming effect of SDR under cloud-sky condition ($`r cloudtrend`\,\%/y$), shows that clouds can also attenuate the amount of SDR reaching the ground, and changes on their occurrences can effect the net energy balance.
Because we have no adequate data to investigate the long term changes of cloud formation in the region, we can not infer if the SDR trend we observe under cloud-sky is biased by changes in clouds and local climate.
The observed brightening of SDR over Thessaloniki; is dependent on SZA (higher SZA, have stronger brightening trends), and also, with the seasons of the year, where Winter and Autumn show a significant strong trend, in contrast to Spring and Summer.
<!-- to conclutions? -->
Using CUSUMs of the trends for all- and clear-skies, we observed periods where the CUSUMs remain relative steady,
<!-- roughly around the middle of the study dataset,--> with a steep decline previous to that, and a steep incline after. This is an indication that the whole brightening phenomenon does not follow a smooth development over time.

>>>>

In order to investigate further the features of the CUSUMs, we created another set of CUSUM plots by subtracting the corresponding long term trend from the SDR anomaly data, prior to the CUSUM calculation (Figure\ \ref{fig:cusumnotrendmonthly}).
With this approach periods when the CUSUMs diverge from zero can be interpreted as a systematic variation of SDR from the climatological mean. 
When the CUSUM is increasing, the added values are above the climatological values of the SDR trend and vice versa.
Overall, for all- and clear-sky conditions (Figures\ \ref{fig:cusumnotrendmonthly-1} and\ \ref{fig:cusumnotrendmonthly-2}) we observe periods when the anomalies diverge from the climatological value, each lasting for several years.
The pattern in both datasets is very similar, suggesting prevalence in clear skies over Thessaloniki.
It is interesting that in the period 1993 - 2016 the anomalies have a high variability around zero, while after 2016, the range of the variability is decreased to about one third of the prior period.
For cloudy-sky conditions (Figure\ \ref{fig:cusumnotrendmonthly-3}) the period 1997 - 2008 is dominated by positive CUSUMs, suggesting a reduced effect of clouds on SDR.
From 1997 to mid-2000s CUSUMs are increasing, likely due to a continuous decrease in the optical thickness of clouds, followed by a period of rapid increase (within 3 years) in cloud optical thickness lasting up to 2008.
The following stable period spans for about 15 years up to 2021 when CUSUMs start increasing again.

@@

Similarly to the long term trends discussed above, we have calculated the trend of the anomalies for the three different sky conditions, and for each season of the year, using the corresponding mean monthly values
(Figure\ \ref{fig:seasonalALL} and Table\ \ref{tab:trendseasontable}).
For all-sky conditions the trend in SDR in winter is the largest
($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$),
followed by the trend in autumn
($`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$,
a value close to the long term trend) both statistically significant above the $99\,\%$ confidence level.
In spring and summer, the trends are much smaller and of lesser statistical significance.
These seasonal differences indicate a possible relation of the trends in SDR to trends of clouds during winter and autumn.
For clear-skies, the trend in winter is $`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear sky"], 2)`\,\%/y$, larger than for all-skies ($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$), which is another indication of a decreasing trend in cloud optical thickness.
Moreover, the trends under clear- and cloudy-sky conditions are almost complementary to each other, particularly for winter and autumn, where the signal is stronger.
During spring and summer the statistical significance is very low and the actual trend too small for a meaningful comparison.

```{r seasonalALL, echo=F, fig.cap="Linear trends (black lines) of monthly mean SDR season (columns of graphs) for the three sky conditions (rows of graphs). The vertical scale is common for all graphs.", out.width="100%"}
knitr::include_graphics("./images/SeasonalTrendsTogether3-2.png")
```

@@

We have demonstrated that in the period
`r paste(year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)))`, there is a brightening trend of SDR for all-sky conditions ($`r alltrend`\,\%/y$).
On a previous study [@Bais2013], for the period 1993 - 2011 the trend was found to be $0.33\,\%/y$.
The increase of this trend, indicates that the phenomenon and probably the causes, are still occurring.
Accordingly, we found a similar trend for clear-sky conditions ($`r cleartrend`\,\%/y$) that further supports that the brightening can be independent of the cloud occurrences, although, cloud formation can also attenuate SDR.
We have calculated the effect of TSI variability to be a negligible factor in SDR variability.
Unfortunately, there is no available data for this period for the aerosols, in order to directly estimate their effect on SDR, which is the most probable main cause.
The attenuation of SDR by aerosols over Europe have been proposed as major factor by @Wild2021.
The dimming effect of SDR under cloud-sky condition ($`r cloudtrend`\,\%/y$), shows that clouds can also attenuate the amount of SDR reaching the ground, and changes on their occurrences can effect the net energy balance.
Because we have no adequate data to investigate the long term changes of cloud formation in the region, we can not infer if the SDR trend we observe under cloud-sky is biased by changes in clouds and local climate.
The observed brightening of SDR over Thessaloniki; is dependent on SZA (higher SZA, have stronger brightening trends), and also, with the seasons of the year, where Winter and Autumn show a significant strong trend, in contrast to Spring and Summer.
<!-- to conclutions? -->
Using CUSUMs of the trends for all- and clear-skies, we observed periods where the CUSUMs remain relative steady,
<!-- roughly around the middle of the study dataset,--> with a steep decline previous to that, and a steep incline after. This is an indication that the whole brightening phenomenon does not follow a smooth development over time.

### 2023-11-04 MDPI_submition.tex
Commit 2023-11-04 14:14

In order to investigate further the features of the CUSUMs, we created
another set of CUSUM plots by subtracting the corresponding long term
trend from the SDR anomaly data, prior to the CUSUM calculation
(Figure~\ref{fig:cusumnotrendmonthly}). With this approach periods when
the CUSUMs diverge from zero can be interpreted as a systematic
variation of SDR from the climatological mean. When the CUSUM is
increasing, the added values are above the climatological values of the
SDR trend and vice versa. Overall, for all- and clear-sky conditions
(Figures~\ref{fig:cusumnotrendmonthly-1}
and~\ref{fig:cusumnotrendmonthly-2}) we observe periods when the
anomalies diverge from the climatological value, each lasting for
several years. The pattern for the all-sky and clear-sky conditions is
very similar, suggesting prevalence in clear skies over Thessaloniki. It
is interesting that in the period 1993 - 2016 the anomalies of all-sky
and clear-sky, have a high variability around zero. After 2016, the
range of the variability is decreased at about one third of the prior
period. For cloud-sky conditions
(Figure~\ref{fig:cusumnotrendmonthly-3}) the period 1997 - 2008 is
dominated by positive CUSUMs suggesting a reduced effect of clouds on
SDR. From 1997 to mid-2000s CUSUMs are increasing, likely due to a
continuous decrease in the thickness of clouds, followed by a period of
rapid increase (within 3 years) in cloud optical thickness lasting up to
2008. The following stable period spans for about 15 years up to 2021
when CUSUMs start increasing again.

@@

Similarly to the long term trends discussed above, we have calculated
the trend of the anomalies for the three different sky conditions, and
for each season of the year, using the corresponding mean monthly values
(Figure~\ref{fig:seasonalALL} and Table~\ref{tab:trendseasontable}). For
all-sky conditions the trend in SDR in winter is the largest
(\(0.69\,\%/y\)), followed by the trend in autumn (\(0.43\,\%/y\), a
value close to the long term trend) both statistically significant above
the \(99\,\%\) confidence level. In spring and summer, the trends are
much smaller and of lesser statistical significance. These seasonal
differences indicate a possible relation of the trends in SDR to trends
of clouds during winter and autumn. For clear-skies, the trend in winter
is \(0.83\,\%/y\), larger than for all-skies (\(0.69\,\%/y\)), which is
another indication of increasing cloud thickness. Moreover, the trends
under clear- and cloud-sky conditions are almost complementary to each
other, particularly for winter and autumn, where the signal is stronger.
During spring and summer the statistical significance is very low and
the actual trend too small for a meaningful comparison.

@@

\caption{SDR trends by season of the year for the three sky conditions. Each point is a monthly mean calculated from daily means. The vertical scale is common for all graphs, each column of graphs corresponds to a sky condition and each row to a season of the year.}\label{fig:seasonalALL}
\end{figure}

@@

We demonstrate that in the period 1993 - 2023, there is a brightening
trend of SDR for all-sky conditions (\(0.38\,\%/y\)). On a previous
study \citep{Bais2013}, for the period 1993 - 2011 the trend was found
to be \(0.33\,\%/y\). The increase of this trend, indicates that the
phenomenon and probably the causes, are still occurring. Accordingly, we
found a similar trend for clear-sky conditions (\(0.35\,\%/y\)) that
further supports that the brightening can be independent of the cloud
occurrences, although, cloud formation can also attenuate SDR. We have
calculated the effect of TSI variability to be a negligible factor in
SDR variability. Unfortunately, there is no available data for this
period for the aerosols, in order to directly estimate their effect on
SDR, which is the most probable main cause. The attenuation of SDR by
aerosols over Europe have been proposed as major factor by
\citet{Wild2021}. The dimming effect of SDR under cloud-sky condition
(\(-0.28\,\%/y\)), shows that clouds can also attenuate the amount of
SDR reaching the ground, and changes on their occurrences can effect the
net energy balance. Because we have no adequate data to investigate the
long term changes of cloud formation in the region, we can not infer if
the SDR trend we observe under cloud-sky is biased by changes in clouds
and local climate. The observed brightening of SDR over Thessaloniki; is
dependent on SZA (higher SZA, have stronger brightening trends), and
also, with the seasons of the year, where Winter and Autumn show a
significant strong trend, in contrast to Spring and Summer. Using CUSUMs
of the trends for all- and clear-skies, we observed periods where the
CUSUMs remain relative steady, with a steep decline previous to that,
and a steep incline after. This is an indication that the whole
brightening phenomenon does not follow a smooth development over time.

>>>>

In order to investigate further the features of the CUSUMs, we created
another set of CUSUM plots by subtracting the corresponding long term
trend from the SDR anomaly data, prior to the CUSUM calculation
(Figure~\ref{fig:cusumnotrendmonthly}). With this approach periods when
the CUSUMs diverge from zero can be interpreted as a systematic
variation of SDR from the climatological mean. When the CUSUM is
increasing, the added values are above the climatological values of the
SDR trend and vice versa. Overall, for all- and clear-sky conditions
(Figures~\ref{fig:cusumnotrendmonthly-1}
and~\ref{fig:cusumnotrendmonthly-2}) we observe periods when the
anomalies diverge from the climatological value, each lasting for
several years. The pattern in both datasets is very similar, suggesting
prevalence in clear skies over Thessaloniki. It is interesting that in
the period 1993 - 2016 the anomalies have a high variability around
zero, while after 2016, the range of the variability is decreased to
about one third of the prior period. For cloudy-sky conditions
(Figure~\ref{fig:cusumnotrendmonthly-3}) the period 1997 - 2008 is
dominated by positive CUSUMs, suggesting a reduced effect of clouds on
SDR. From 1997 to mid-2000s CUSUMs are increasing, likely due to a
continuous decrease in the optical thickness of clouds, followed by a
period of rapid increase (within 3 years) in cloud optical thickness
lasting up to 2008. The following stable period spans for about 15 years
up to 2021 when CUSUMs start increasing again.

@@

Similarly to the long term trends discussed above, we have calculated
the trend of the anomalies for the three different sky conditions, and
for each season of the year, using the corresponding mean monthly values
(Figure~\ref{fig:seasonalALL} and Table~\ref{tab:trendseasontable}). For
all-sky conditions the trend in SDR in winter is the largest
(\(0.69\,\%/y\)), followed by the trend in autumn (\(0.43\,\%/y\), a
value close to the long term trend) both statistically significant above
the \(99\,\%\) confidence level. In spring and summer, the trends are
much smaller and of lesser statistical significance. These seasonal
differences indicate a possible relation of the trends in SDR to trends
of clouds during winter and autumn. For clear-skies, the trend in winter
is \(0.83\,\%/y\), larger than for all-skies (\(0.69\,\%/y\)), which is
another indication of a decreasing trend in cloud optical thickness.
Moreover, the trends under clear- and cloudy-sky conditions are almost
complementary to each other, particularly for winter and autumn, where
the signal is stronger. During spring and summer the statistical
significance is very low and the actual trend too small for a meaningful
comparison.

@@

\caption{Linear trends (black lines) of monthly mean SDR season (columns of graphs) for the three sky conditions (rows of graphs). The vertical scale is common for all graphs.}\label{fig:seasonalALL}
\end{figure}

@@

We have demonstrated that in the period 1993 - 2023, there is a
brightening trend of SDR for all-sky conditions (\(0.38\,\%/y\)). On a
previous study \citep{Bais2013}, for the period 1993 - 2011 the trend
was found to be \(0.33\,\%/y\). The increase of this trend, indicates
that the phenomenon and probably the causes, are still occurring.
Accordingly, we found a similar trend for clear-sky conditions
(\(0.35\,\%/y\)) that further supports that the brightening can be
independent of the cloud occurrences, although, cloud formation can also
attenuate SDR. We have calculated the effect of TSI variability to be a
negligible factor in SDR variability. Unfortunately, there is no
available data for this period for the aerosols, in order to directly
estimate their effect on SDR, which is the most probable main cause. The
attenuation of SDR by aerosols over Europe have been proposed as major
factor by \citet{Wild2021}. The dimming effect of SDR under cloud-sky
condition (\(-0.28\,\%/y\)), shows that clouds can also attenuate the
amount of SDR reaching the ground, and changes on their occurrences can
effect the net energy balance. Because we have no adequate data to
investigate the long term changes of cloud formation in the region, we
can not infer if the SDR trend we observe under cloud-sky is biased by
changes in clouds and local climate. The observed brightening of SDR
over Thessaloniki; is dependent on SZA (higher SZA, have stronger
brightening trends), and also, with the seasons of the year, where
Winter and Autumn show a significant strong trend, in contrast to Spring
and Summer. Using CUSUMs of the trends for all- and clear-skies, we
observed periods where the CUSUMs remain relative steady, with a steep
decline previous to that, and a steep incline after. This is an
indication that the whole brightening phenomenon does not follow a
smooth development over time.
